Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees

### 2023-01-22 script/manuscript_MEE.Rmd
patched up writing with properly run blast classification!

aa <- prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)
bb <- prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)
cc <- prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)
dd <-  prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
ee <- prcnt_detect(ericaceae_detected, Lprop = 0.001, Hprop = 0.1)

@@

rm(mertensia_samples, blst, blst_aster, blst_aster1,
   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
   blst_mertensia, aster_samples, ericaceae_samples, blst_ericoid)
```

@@

54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing. The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`). Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`. The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))`.

**APPENDIX X Reads Per Loci**.

@@

sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs'] )
sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs'] )
fa_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs'] )
```

The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources). A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix XX). Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches. Each of the top ten taxa which were identified by BLAST from the aforementioned process composed `r sum(pp_seqs_t$Samples_Classified)` distinct records (CONFUSING WORDS), of these `r seqs_sp`% of the reads were classified to species `r sp_total_seqs`% which represent of the total classified sequences, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of the classified reads, and `r seqs_fa`% of the records were classified to genus `r fa_total_seqs`% of total sequences.

*Sequences which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps. These were manually assigned to a variety of ranks, occasionally to genus #COUNTS# and family #COUNTS#, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new). * <- went back on this as new function captures most of it, but don't throw out.

To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results. The pollen samples contained three morphotypes which could readily be identified via microscopy. Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae. Boraginaceae grains `r toString(aa)`.  Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`. Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.

```{r clean sequence numbers from environment}
rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, 
    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa)
```

@@

To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads. In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g. if both species of *Mertensia*, or one species and read only classified to genus were present in a sample, the reads were summed. The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.

The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated. This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021). There was strong evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).

@@

e <- blst %>% 
  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae')) %>% 
  select(sample.id, Genus) %>% 
  mutate(Present = 1) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  pivot_wider(names_from = Genus, values_from = Present) %>% 
  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
  column_to_rownames('sample.id')

@@

To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard,  $\alpha$ index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha). Numerous taxa from the family Ranunculaceae (*Caltha* L. sp., *Thalictrum* L. spp., *Trollius* L. sp., *Aquilegia* L. spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz. were recorded. A similar relationship was observed in the Hydrophyllaceae with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L. species were present.

@@

blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
  rename(sample.id = Sample, Genus = GENUS, Family = FAMILY) %>% 
  distinct(sample.id)

blst <- read.csv('../data/blast_seqs_classified.csv') %>% 
  mutate(across(where(is.character), ~ replace_na(., 'UNKNOWN'))) %>% 
  arrange(sample.id, -Prcnt_seqs) %>% 
  inner_join(., corb, by = 'sample.id')

@@

tb <- blst %>% 
  left_join(., blst_cnt, by = 'species') %>% 
  group_by(species, name) %>% 
  mutate(Mu_Seqs = sum(Prcnt_seqs)/n,
         species = str_remove(species, 'Bombus '))  %>% 
  distinct(species, name, .keep_all = T) %>% 
  ungroup(name) %>% 
  slice_max(Mu_Seqs, n = 10) %>% 
  select(plant.species = name, species, n, Mu_Seqs, Genus, Family) %>% 
  mutate(rank = 1:n()) %>% 
  ungroup() %>% 
  mutate(Family = case_when(
    Genus %in% c('Trollius', 'Caltha', 'Aquilegia', 'Delphinium') ~ 'Ranunculaceae',
    Genus == 'Hydrophyllum' ~ 'Hydrophyllaceae',
    Genus == 'Epilobium' ~ 'Onagraceae',
    Genus == 'Viola' ~ 'Violaceae',
    Genus == 'Salix' ~ 'Saliceae',
    Genus == 'Cynoglossum' ~ 'Boraginaceae', 
    Genus == c('Arenaria', 'Odontostemma') ~ 'Caryophyllaceae',
    Genus == 'Parnassia' ~ 'Celastraceae',
    Genus %in% c('Erigeron', 'Symphyotrichum') ~ 'Asteraceae',
    TRUE ~ Family 
  ))

@@

ggplot(data=correct, aes(x=Observation, y=value,
                         group = Bombus.Species, color = Bombus.Species)) +
  geom_line(linewidth = 1.25, linetype = 'dotted')+
  geom_point(size = 3, shape = 23, aes(fill = Bombus.Species)) +
  scale_fill_manual('Bee species', values = bee_cols) +
  scale_color_manual('Bee species', values = bee_cols) +
  theme_bw() +
  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
       y = 'Shared Observations', x= NULL) + 
  theme(panel.grid.major.x = element_blank(),
        panel.border = element_blank())

@@

@@

... many of our results indicate foraging on *Viola* L. spp, zygomorphic flowers with architecture which would require subtle handling and strength to reach the pollen and nectar loads... .. Or the *Epilobium* L. spp. results likely indicating that a species of *Chamerion* Seg. such as *C. angustifolium* (L.) Scop.  or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.

@@

We anticipate that many of the complications which we faced, using opportunistically collected pollen loads and the first implementation of this method may readily be overcome. It seems apparent that we had issues detecting pollen from several genera of plants, based upon these and other observational studies most likely *Vicia* L., *Lathyrus* L., and *Frasera* (@inouye1980effect, @pleasants1980competition); this is most likely related to user error in obtaining high quality DNA during the plant reference library generation period.  **(REED SHOULD HUNT FOR LOCI RETURNS FROM BLAST LIKE HE DID POLLEN)**. Additional complications seem to relate to the presence of closely related false positives, e.g. frequent classifications of sequences as *Trollius* L.,  *Caltha* L., and *Thalictrum* L. alongside a more common species in the family, e.g. *Delphinium.* L. Many of our errors are known to us and multiple mnemonics are in *APPENDIX XX* to assist others in future attempts to achieve better results.  However, the line between false and negative positives may be blurred in some of these instances and warrant further work, for example **Ericaceae pollen grains were observed in a number of samples in trace quantities...**.  We used only a relatively few amount of species records for both the spatial and temporal modelling, we feel this was an appropriate measure to show that these methods would work with data sets which were readily available without considerable data 'scraping'. We anticipate that the number of records available to all researchers have grown rapidly since the inception of our modelling (2019) to date, largely as a result of herbarium digitization efforts, and that speak of a massive 'one stop shop' of an integrated database is just over the horizon (@feng2022review, @soltis2017digitization).
  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area. However, we had difficulties in evaluating our predictions in an operational context. We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local). While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult. We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data. Further given the, size of the minimum spanning tree (AREA???) which we extracted points to, a formal floristic inventory would still be a time intensive process. Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding. We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts. 
  Difficulties in temporal models related to variability in drivers of flowering phenology.

@@

Results from palynological analyses show that several species of bee show near perfect fidelity to the genus Hydrophyllum on a per visit basis.

>>>>

aa <- prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)
bb <- prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)
cc <- prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)
dd <- prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
ee <- prcnt_detect(ericaceae_detected, Lprop = 0.001, Hprop = 0.1)
ff <- prcnt_detect(aster_detected, Lprop = 0.001, Hprop = 0.01)

@@

rm(mertensia_samples, blst_aster, blst_aster1,
   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
   blst_mertensia, aster_samples, ericaceae_samples, blst_ericoid)
```

@@

54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing. The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`). Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`. The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))`. **APPENDIX X Reads Per Loci**.

... with samples 22, and 41 both having < 100 reads indicates, virtual failure of these records (REMOVE from analyses).

@@

sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs'] )
sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs'] )
fa_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs'] )

byhand <- blst %>% 
  filter(TAXON_NEW != TAXON_NEW2)

j <- toString(paste0(names(table(byhand$Condition)), ": ",
                     as.vector (table(byhand$Condition))))
gen <- byhand[byhand$TAXON_NEW2==c('Salix'),]
spp <- byhand[byhand$TAXON_NEW2==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
```

The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources). A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix XX). Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches. Each of the top ten taxa which were identified by BLAST from the aforementioned process composed `r sum(pp_seqs_t$Samples_Classified)` distinct records, of these `r seqs_sp`% of the reads were classified to species `r sp_total_seqs`% representing of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to genus `r fa_total_seqs`%  of all total sequences.

`r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps. These were all assigned via post-processing conditions (`r j`). These were manually assigned to a variety of ranks, occasionally to genus `r nrow(gen)` and species `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).

To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results. The pollen samples contained three morphotypes which could readily be identified via microscopy. Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae. Boraginaceae grains `r toString(aa)`.  Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`. Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.

```{r clean sequence numbers from environment}
rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
```

@@

To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads. In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g. if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed. The total percentage of the ten most abundant grains per sample were then were then relativize to constitute the entire sample.

The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated. This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021). There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).

@@

e <- blst %>% 
  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
                       'Celastraceae')) %>% 
  select(sample.id, Genus) %>% 
  mutate(Present = 1) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  pivot_wider(names_from = Genus, values_from = Present) %>% 
  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
  column_to_rownames('sample.id')

@@

To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, $\alpha$ index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha). Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L. sp., *Thalictrum* L. spp., *Trollius* L. sp., *Aquilegia* L. spp.), had ... $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz. were recorded. A similar relationship was observed in the Hydrophyllaceae R.Br. with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L. species were present.

@@

blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
  rename(sample.id = Sample, Genus = GENUS, Family = FAMILY) %>% 
  arrange(sample.id, -Prcnt_seqs) %>% 
  inner_join(., corb, by = 'sample.id')

@@

tb <- blst %>% 
  left_join(., blst_cnt, by = 'species') %>% 
  group_by(species, TAXON_NEW2) %>% 
  mutate(Mu_Seqs = sum(Prcnt_seqs)/n,
         species = str_remove(species, 'Bombus '))  %>% 
  distinct(species, TAXON_NEW2, .keep_all = T) %>% 
  ungroup(TAXON_NEW2) %>% 
  slice_max(Mu_Seqs, n = 10) %>% 
  select(plant.species = TAXON_NEW2, species, n, Mu_Seqs, Genus, Family) %>% 
  mutate(rank = 1:n()) %>% 
  ungroup() %>% 
  mutate(Family = case_when(
    Genus %in% c('Trollius', 'Caltha', 'Aquilegia', 'Delphinium') ~ 'Ranunculaceae',
    Genus == 'Hydrophyllum' ~ 'Hydrophyllaceae',
    Genus == 'Epilobium' ~ 'Onagraceae',
    Genus == 'Viola' ~ 'Violaceae',
    Genus == 'Salix' ~ 'Saliceae',
    Genus == 'Cynoglossum' ~ 'Boraginaceae', 
    Genus == c('Arenaria', 'Odontostemma') ~ 'Caryophyllaceae',
    Genus == 'Parnassia' ~ 'Celastraceae',
    Genus %in% c('Erigeron', 'Symphyotrichum') ~ 'Asteraceae',
    TRUE ~ Family 
  ))

@@

ggplot(data=correct, aes(x=Observation, y=value,
                         group = Bombus.Species, color = Bombus.Species)) +
  geom_line(linewidth = 1.25, linetype = 'dotted', alpha = 0.3)+
  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
              height = 0.05, width = 0.05) +
  scale_fill_manual('Bee species', values = bee_cols) +
  scale_color_manual('Bee species', values = bee_cols) +
  theme_bw() +
  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
       y = 'Shared Observations', x= NULL) + 
  theme(panel.grid.major.x = element_blank(),
        panel.border = element_blank())

@@

```{r}

blst_final <- blst %>% 
  mutate(TAXON_EXPERTS = case_when(
      TAXON_NEW2 %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
      'Cynoglossum officinale') ~ 'Mertensia',
      TAXON_NEW2 %in% c('Caltha leptosepala', 'Trollius europaeus', 'Trollius') ~ 'Delphinium',
    # 'Thalictrum thalictroides' ~ 'Delphinium',
      TAXON_NEW2 == 'Erigeron' ~ 'Erigeron grandiflorus', 
      TAXON_NEW2 == 'Scabrethia scabra' ~ 'Wyethia arizonica',
      TAXON_NEW2 == 'Nemophila breviflora' ~ 'Hydrophyllum',
      TAXON_NEW2 == 'Erigeron grandiflorus' ~ 'Erigeron grandiflorus',
      TAXON_NEW2 == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
      TAXON_NEW2 == 'Paxistima myrsinites' ~ 'Parnassia palustris', # original was probably right !!! shows the consequence of bad time estimates.#, 
    # TRUE ~ as.character(TAXON_NEW2)
  )) %>% 
  select(TAXON, TAXON_NEW2, TAXON_EXPERTS)

# no TARAXACUM OFICINALE, shows a limitation of the data sources.

Situations where SDM's led to incorrect results at the species level are evident with classification to *Scabrethia Scabra* (Hooker) W.A. Weber, this match almost certainly representing *Wyethia arizonica* A. Gray (@weber1998new), a taxon known to be visited by Queen bee's via our floral observations.

@@

... many of our results indicate foraging on *Viola* L. spp, zygomorphic flowers with architecture which would require subtle handling and strength to reach the pollen and nectar loads... (IS FORREST PAPER WORTH CITING ? IS THIS EVEN WORTH HAVING?)

An expected inaccuracy of the classification scheme is in genus level placements, e.g. were *Epilobium* L. (Onagraceae Juss.) spp. were classified. However, given the small size of their flowers in the study area, these results more likely indicate that a species of *Chamaenerion* Seg. (a segregate genus) such as *C. angustifolium* (L.) Scop.  or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.

Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed reclassifications of the contents of pollen grains...

<final integrative results here>

@@

  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area. However, we had difficulties in evaluating our predictions in an operational context. We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local). While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult. We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data. Further given the, size of the minimum spanning tree (AREA???) which we extracted points to, a formal floristic inventory would still be a time intensive process. Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding. We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts. 
  Difficulties in temporal models related to variability in drivers of flowering phenology.

@@

Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Hydrophyllum* on a per visit basis... General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persit with *Bombus* studies...
